---
title: "Homework 1"
author: "Sumit Saurabh"
date: "Due 09-11-2023"
output:
  html_document: default
  pdf_document: default
subtitle: DSCI 101
editor_options:
  chunk_output_type: console
---

Homework should be written as an R Markdown file and the knitted output should be submitted. Homework should be turned in on Sakai.

1.  (4pts) Set up a github account and create a repository called DSCI101. Clone this repo into RStudio: <https://github.com/menawhalen/DSCI_101_F23>

2.  (3pts) Create a numeric vector containing the values 5, 10, 15, 20, and 25. Calculate the mean of this vector and assign it to a variable called `mean_value`. Print the `mean_value`

```{r ex2}
vec1<-c(5,10, 15,20, 25)
mean_value<-mean(vec1)
print(mean_value)
```

3.  (3pts) Create a data frame with the following information about three of your friends:

-   Name (character)
-   Age (numeric)
-   Favorite Color (character)

Make sure to include at least three rows of data. Print the entire data frame. Name it whatever you like.

```{r ex3}
name <-c("Jenn", "Dawn", "Craig")
age <- c(23, 24, 50)
fav_color<-c("blue", "white", "red")
df<-data.frame(name, age, fav_color)
print(df)
```

4.  (3pts) Create two numeric vectors: `vector1` with values 1, 2, 3, and `vector2` with values 4, 5, 6. Perform the following operations:

-   Add `vector1` and `vector2` and store the result in a new vector called `sum_vector.`
-   Subtract `vector2` from `vector1` and store the result in a new vector called `diff_vector.`
-   Multiply `vector1` and `vector2` element-wise and store the result in a new vector called `prod_vector.`
-   Divide `vector1` by `vector2` element-wise and store the result in a new vector called `quot_vector.`

Print all four result vectors.

```{r ex4}
Vector1 <- c(1,2,3)
Vector2 <- c(4,5,6)
sum_vector <- c(Vector1 + Vector2)
diff_vector <- c(Vector2-Vector1)
prod_vector <- c(Vector1*Vector2)
quot_vector <- c(Vector1/Vector2)
info <- c(print(sum), print (diff_vector), print(prod_vector), print(quot_vector))

```

5.  (12pts) Using the `top_colleges_2022` dataset answer the following questions.

```{=html}
<!-- -->
```
a)  Read in the csv file and name the variable `college`. This should be simple if question 1 has been completed as the github will download the data folder.

```{r ex5a}
library(readr)
library(tidyverse)
college <- read_csv("data/top_colleges_2022.csv")

```

b)  How many colleges in this data set are in Illinois.

```{r ex5b}
LUCcolleges<-filter(college, state...4 == "IL")
countLUC<-count(LUCcolleges)
countLUC<-count(LUCcolleges)
print(countLUC)

```

c)  Find the undergraduate student population of LUC. How man colleges (in this dataset) have a larger undergraduate student population than LUC?

```{r ex5c}
LUC2 <- filter (college, organizationName == "Loyola University Chicago")
print(LUC2$studentPopulation)

```
